---
title: 'GSoC Week 1:  June/06-June/09'
author: "Erick Calderon-Morales"
text: "Mentor "
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 2
---

# Remote execution 

+ __For the remote execution task I was able to create a tunnel following the PEcAn documentation, however this only worked when I used the tmp folder inside the SCC.__

```{bash eval = FALSE}
# Remote access ----------------------------------------------------------------
# Step 1: Create tunnel in BASH
ssh -n -N -f -o ControlMaster=yes -S /tmp/ecalder1_gsoc_1 ecalder1@geo.bu.edu
```

```{r eval = FALSE}
# Step 2: access tunnel
scc_host <- list(name = "geo.bu.edu", tunnel = "/tmp/ecalder1_gsoc_1")
 
# Step 3: Test that is working
PEcAn.remote::remote.execute.cmd(host = scc_host,
                                  cmd = "echo", args = "Hello world")
```


+ __I tried to create the tunnel using a different path but I was unsuccessful and I got the error:__

`unix_listener`: __cannot bind to path /usr2/collab/ecalder1/tunnel/ecalder1_gsoc_1.QVMno7DENxYuUiFl: No such file or directory__


```{bash eval = FALSE}
ssh -n -N -f -o ControlMaster=yes -S /usr2/collab/ecalder1/tunnel/ecalder1_gsoc_1 ecalder1@geo.bu.edu
 
ssh -nNf -o ControlMaster=yes -S /projectnb/dietzelab/ecalder1/tunnel/ecalder1_gsoc_1 ecalder1@geo.bu.edu
```


# Meta-Analysis and Ensemble 

+ __I created a test XML file for running and ensemble and a meta-analysis locally. For this I tried to combine Alexis' pecan.CONFIGS.xml file with this other [XML file](https://gist.github.com/dlebauer/7a02538ac06ce4bab6354f1d226555dc).__


## Test XML file 

```{xml}
<?xml version="1.0"?>
<pecan>
  
 <info>
  <notes>Daily Forecast SIPNET Site</notes>
  <userid>1000012038</userid>
  <username>ecalder1</username>
  <date>2022/06/09</date>
 </info>
 
 <database>
    <bety>
      <user>bety</user>
      <password>bety</password>
      <host>postgres</host>
      <dbname>bety</dbname>
      <driver>PostgreSQL</driver>
      <port>5432</port>
      <write>true</write>
    </bety>
    <dbfiles>/home/carya/gsoc/pecan.data/dbfiles</dbfiles>
 </database>
 
 <host>
  <name>geo.bu.edu</name>
  <user>ecalder1</user>
  <outdir>/home/carya/gsoc/ecalder1/Jobs/out</outdir>
 </host>
 
 
 <pfts>
  <pft>
   <name>temperate.deciduous.HPDA</name>
   <constants>
    <num>1</num>
   </constants>
   <posteriorid>1000022311</posteriorid>
   <outdir>/home/carya/gsoc/output/PEcAn_1000010530/pft/temperate.deciduous.HPDA</outdir>
  </pft>
 </pfts>
 
 <meta.analysis>
  <iter>3000</iter>
  <random.effects>
   <on>FALSE</on>
   <use_ghs>FALSE</use_ghs>
  </random.effects>
  <threshold>1.2</threshold>
  <update>AUTO</update>
 </meta.analysis>
 
 <ensemble>
    <size>100</size>
    <variable>GPP</variable>
    <start.year>2020</start.year>
    <end.year>2020</end.year>
    <samplingspace>
     <parameters>
      <method>uniform</method>
     </parameters>
     
    <met>
     <method>sampling</method>
    </met>
    <soil>
     <parent>parameters</parent>
    </soil>
    <vegetation>
     <parent>soil</parent>
    </vegetation>
   </samplingspace>
  <ensemble.id>1000033746</ensemble.id>
 </ensemble>

</pecan>
```


+ __With this test XML I was able to run the meta analysis however I couldn't run the ensemble__

```{r message=FALSE, warning=FALSE}
# Load packages ----------------------------------------------------------------
library(PEcAn.all)
library(PEcAn.utils)
library(PEcAn.workflow)
setwd("/home/carya")
getwd()
```


```{r message=TRUE, warning=TRUE}
# Read settings file -----------------------------------------------------------
settings <- read.settings("/home/carya/gsoc/pecan.xml")
```

```{r error = TRUE}
# Test get trait data function -------------------------------------------------
get.trait.data(settings$pfts, settings$model$type, 
                                settings$database$dbfiles, 
                                settings$database$bety, 
                                settings$meta.analysis$update)
```
`Error: Column `name` not found in `.data` Run `rlang::last_error()` to see where the error occurred.`

## Meta Analysis

```{r error = TRUE}
# Run meta analysis ------------------------------------------------------------
PEcAn.MA::runModule.run.meta.analysis(settings)
```
## Ensemble

```{r error =TRUE}
# Run ensemble -----------------------------------------------------------------
dir.create(settings$pfts$pft$outdir, recursive = TRUE, showWarnings = FALSE)
dir.create(settings$database$dbfiles, recursive = TRUE, showWarnings = FALSE)


runModule.run.ensemble.analysis(settings, TRUE)

```


+ __I tried to run the xml test file remotely but I was unsuccessful__ 
```{r error = TRUE}
PEcAn.remote::runModule.start.model.runs(settings = settings)
```



## Need guidence on...

+ __Identify what errors I am making on the test XML file that I created for being able to run the Ensemble__
+ __How to run XML files on the SCC__


